\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R

### 2014-01-22 papers/jss/article.Rnw
quick round over Summary
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@818 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{Concluding remarks}
\label{sec:summary}
% TODO(mstokely): Get cibona approval for these two sentences before
% publishing.
Schema-less text formats such as CSV and JSON will continue to be
widely used in many contexts, but we hope that the availability of
\pkg{RProtoBuf} makes it easy for many mixed-language data analysis
pipelines to embrace schemas such as Protocol Buffers for type-safe
and performant data serialization between applications.

\pkg{RProtoBuf} has been heavily used inside Google for the past three
years by statisticians and software engineers.  At the time of this
writing there are more than XXX 30-day active users of RProtoBuf using
it to read data from and otherwise interact with other distributed
systems written in C++, Java, Python, and other languages.
\\

\emph{Other Approaches}
\\

\pkg{RProtoBuf} is quite flexible and easy to use for interactive use,
but it is not designed for efficient high-speed manipulation of large
numbers of protocol buffers once they have been read into R.  For
example, taking a list of 100,000 Protocol Buffers, extracting a named
field from each one, and computing an aggregate statistic on those
values would be relatively slow with RProtoBuf.  Mechanisms to address
such use cases are under investigation for possible incorporation into
future releases of RProtoBuf, but currently, the package relies on
other database systems to provide query and aggregation semantics
before the resulting protocol buffers are read into R.  Inside Google,
the Dremel query system \citep{dremel} is often employed in this role
in conjunction with \pkg{RProtoBuf}.

@@

The first versions of \CRANpkg{RProtoBuf} were written during 2009-2010.
Very significant contributions, both in code and design, were made by
Romain Fran\c{c}ois whose continued influence on design and code is
greatly appreciated. Several features of the package are influenced
by the design of the \CRANpkg{rJava} package by Simon Urbanek
The user-defined table mechanism, implemented by Duncan Temple Lang for the
purpose of the \pkg{RObjectTables} package, allows for the dynamic symbol lookup.
Kenton Varda was generous with his time in reviewing code and explaining
obscure protocol buffer semantics.  Karl Millar was very
helpful in reviewing code and offering suggestions.  
%The contemporaneous work by Saptarshi Guha on \pkg{RHIPE} was a strong
%initial motivator.

>>>>

\section{Summary}  % DE Simpler title
\label{sec:summary}
% TODO(mstokely): Get cibona approval for these two sentences before
% publishing
Over the past decade, many formats for interoperable
data exchange have become available, each with their unique features,
strengths and weaknesses.  
Text based formats such as CSV and JSON are easy to use, and will likely 
remain popular among statisticians for many years to come. However, in the 
context of increasingly complex analysis stacks and applications involving 
distributed computing as well as mixed language analysis pipelines, choosing a more 
sophisticated data interchange format may reap considerable benefits. 
The Protocol Buffers protocol and librart offers a unique combination of features, performance,
maturity, and forward-compatibility that seems particulary well suited for data-driven 
applications and numerical computing.

%% DE Re-ordering so that we end on RProtoBuf
The \pkg{RProtoBuf} package builds on the Protocol Buffers library, and
extends the R system with the ability to create, read and write Protocol
Buffer message. \pkg{RProtoBuf} has been used extensively inside Google 
for the past three years by statisticians, analysts and software engineers.
At the time of this writing there are more than XXX 30-day active
users of \pkg{RProtoBuf} using it to read data from and otherwise interact
with other distributed systems written in C++, Java, Python, and 
other languages.

As the \pkg{RProtoBuf} package provides users with the ability to generate,
parse and manipulate Protocol Buffer messages in R, it is our hope that this
package will make Protocol Buffers more accessible to the R community, and
thereby makes a small contribution towards better integration between R and
other software systems and applications.

%\emph{Other Approaches}
%
%== JO: I don't really like this section here, it gives the entire paper a bit of a 
%sour aftertaste. Perhaps we can mention performance caveats in the technical
%sections? I think it's nicer to leave it at the above paragraphs.==
%
% DE: Agreed -- commenting out

%% \pkg{RProtoBuf} is quite flexible and easy to use for interactive use,
%% but it is not designed for efficient high-speed manipulation of large
%% numbers of protocol buffers once they have been read into R.  For
%% example, taking a list of 100,000 Protocol Buffers, extracting a named
%% field from each one, and computing an aggregate statistic on those
%% values would be relatively slow with RProtoBuf.  Mechanisms to address
%% such use cases are under investigation for possible incorporation into
%% future releases of RProtoBuf, but currently, the package relies on
%% other database systems to provide query and aggregation semantics
%% before the resulting protocol buffers are read into R.  Inside Google,
%% the Dremel query system \citep{dremel} is often employed in this role
%% in conjunction with \pkg{RProtoBuf}.

@@

The first versions of \CRANpkg{RProtoBuf} were written during 2009-2010.
Very significant contributions, both in code and design, were made by
Romain Fran\c{c}ois whose continued influence on design and code is
greatly appreciated. Several features of the package reflect
the design of the \CRANpkg{rJava} package by Simon Urbanek
The user-defined table mechanism, implemented by Duncan Temple Lang for the
purpose of the \pkg{RObjectTables} package, allows for the dynamic symbol lookup.
Kenton Varda was generous with his time in reviewing code and explaining
obscure protocol buffer semantics.  Karl Millar was very
helpful in reviewing code and offering suggestions.  
%The contemporaneous work by Saptarshi Guha on \pkg{RHIPE} was a strong
%initial motivator.
